Inefficient Concessions and Mediation

### 2021-01-31 draft.tex
change corollary 2 to theorem 6a parameterization; simplify e-g calculations.R

We introduce a two-sided signaling model in which concessions can provide future value to the recipient and/or serve as a signal that provides the recipient with information about the type of player the conceding state is. Using a mechanism design framework, we demonstrate that a mediator can facilitate more efficient concessions by removing uncertainty about the ability of the parties to commit to peace.

@@

\begin{corollary}
	Assume that high types are not patient enough to separate without concessions but that they can separate using concessions when gifts have no material value. If negotiating parties use concessions with future material value that the low-type country can use to reduce the future payoffs of its negotiating partner instead of concessions with no future material value, there are parameters under which the future material value of gifts destroys negotiating partners' ability to separate through concessions, i.e. peace cannot be achieved when there are two high types.
	\label{theorem:matval}
\end{corollary}

\emph{Proof}: An example suffices as proof. Let $T = 10$, $W=8$ and $D = 5$. The assumption that $T > W - D$ is satisfied, and the threshold for high types to separate when concessions do not have future material value is $\de = \frac{8}{10+5} = .5\overline{3}$. Let $\de_h = .6$ and $\de_l = .3$. Assume $p = .2$. By Theorems~\ref{theorem:1} and ~\ref{theorem:2}, both the \emph{no concessions separating equilibrium} and the \emph{concessions separating equilibrium} are possible when gifts have no material value.separation through concessions is possible.

The patience threshold to separate through concessions with efficient gifts () is 
is much higher under these parameters when the concessions carry future material value. The minimum patience level required for separation is $\de_h^1 = \frac{p(W-D-T)}{\frac{\left(1-\de_l\right)p\left(D +	 T\right)}{\left(1-\de_l\right)\left(1 - pT \right) + (1-p) D}} + p (W-T) + (1-p)D + 1 = 2.4$, with a minimum separating concession of $g^1 = \frac{p(1-d_l)(D+T)}{(1-d_l)(1-pT)+(1-p)D} = .\overline{63}$. Since discount factors can be no larger than 1, we have shown that under this set of parameters, the concessions separating equilibrium when concessions have material value cannot be sustained.

$\de_h$ is not involved in any of these calculations, so we can see that for any $\de_h \in [.5\overline{3},1)$.\footnote{Recall that we restrict $p<1$ since there is no imperfect information when only one type is possible.} For this range of $\de_h$, separating through concessions is possible when concessions have no material value but is not possible when they do have future material value. That is, future material value destroys the opportunity for peace. \hfill $\blacksquare$

@@

Thus the constraint for both incentive compatibility and individual rationality is
\begin{equation*}
  \frac{p}{1-\de_h}\left(T + Teg \right) - pg +peg+ \frac{1-p}{1-\de_h}\left(W-D\right) \geq \frac{1}{1-\de_h}\left(W-D\right)
\end{equation*}
This simplifies to
\begin{equation*}
  \frac{1}{1-\de_h}\left(T + Teg \right) - g +eg \geq \frac{1	}{1-\de_h}\left(W-D\right)
\end{equation*}
\begin{equation*}
  T + Teg - (1-\de_h)(1-e)g \geq W-D
\end{equation*}
\begin{equation*}
  T + Teg - (1-\de_h)(1-e)g \geq W-D
\end{equation*}
\begin{equation*}
  T + D - W \geq \left[Te - (1-\de_h)(1-e)\right]g 
\end{equation*}
If $Te < (1-\de_h)(1-e)$, this is a negative lower bound on the size of the concession and thus implies no additional restriction. If $Te = (1-\de_h)(1-e)$, there is again no restriction because $T+D >W$ by assumption. If $Te > (1-\de_h)(1-e)$, the mechanism is both individually rational and incentive compatible for the high type when $\frac{ T + D - W}{Te - (1-\de_h)(1-e)}>g$. \hfill $\blacksquare$\\
\\
\emph{Proof of Theorem~\ref{theorem:medwel}:}\\
An example suffices to prove each part.
	\begin{enumerate}
		\item[(a)] Let $T = 1$, $W=1$ and $D = 1$. The assumption that $T > W - D$ is satisfied, and the threshold for high types to separate through concessions when the future material value of concessions is destroyed is $\de = \frac{1}{1+1} = .5$. Let $\de_h = 0.7$ and $\de_l = .3$ and $p = .9$. By Theorem~\ref{theorem:2}, separation through concessions is possible.

By Theorem~\ref{theorem:csemv}(a), the patience threshold to separate through concessions with efficient concessions is $0.979$ so that the \emph{concessions separating equilibrium with efficient gifts} is not possible while the \emph{concessions separating equilibrium with inefficient gifts} is possible.

Using the mediation mechanism $M$ with $e=0.29$ and $g=12.99$, we satisfy both the low type constraint the $g\geq \frac{\left(1 - \de_l \right)\left(T +D \right)}{p - e\left(p +(1-\de_l)T +W \right)}=9.5890411$ and the high type constraint $g\leq \frac{ T + D - W}{Te - (1-\de_h)(1-e)}=12.987013$. Welfare is 5.0784, compared to 2.82 in the \emph{concessions separating equilibrium with inefficient gifts}.

		\item[(b)] Let $T = 1$, $W=1$ and $D = 1$. The assumption that $T > W - D$ is satisfied, and the threshold for high types to separate through concessions when the future material value of concessions is destroyed is $\de = \frac{1}{1+1} = .5$. Let $\de_h = \frac{13}{16}=0.9375$ and $\de_l = .5$ and $p = .5$. By Theorem~\ref{theorem:2}, separation through concessions is possible.

>>>>

We introduce a two-sided signaling model in which concessions can provide future value to the recipient and/or serve as a signal that provides the recipient with information about the type of player the conceding state is. Using a mechanism design framework, we demonstrate that a mediator can facilitate more efficient concessions by removing uncertainty about the ability of the parties to commit to peace. Perhaps more importantly, the mediator can also sometimes facilitate peace when the negotiating parties cannot achieve it through bilateral interaction.

@@

\begin{corollary}
	Assume that high types are not patient enough to separate without concessions but that they can separate using concessions when gifts have no material value. If negotiating parties use concessions with future material value that the low-type country can use to reduce the future payoffs of its negotiating partner instead of concessions with no future material value, there are parameters under which the future material value of gifts destroys negotiating partners' ability to separate through concessions, i.e. peace cannot be achieved when there are two high types.
	\label{corollary:matval}
\end{corollary}

\emph{Proof}: An example suffices as proof. Let $T = 1$, $W=1$ and $D = 1$. The assumption that $T > W - D$ is satisfied, and the threshold for high types to separate when concessions do not have future material value is $\de = \frac{1}{1+1} = .5$. Let $\de_h = 0.7$ and $\de_l = .3$ and $p = .9$.

By Theorems~\ref{theorem:1} and ~\ref{theorem:2}, both the \emph{no concessions separating equilibrium} and the \emph{concessions separating equilibrium} are possible when gifts have no material value.

The patience threshold to separate through concessions with efficient gifts is much higher under these parameters when the concessions carry future material value: by Theorem~\ref{theorem:csemv}(a), this threshold is $0.979$. Thus the \emph{concessions separating equilibrium with efficient gifts} is \emph{not} possible.

$\de_h$ is not involved in any of these calculations, so we can see that for any $\de_h \in [.5,.979)$, separating through concessions is possible when concessions have no material value but is not possible when they do have future material value. That is, future material value destroys the opportunity for peace. \hfill $\blacksquare$

@@

Thus the constraint for both incentive compatibility and individual rationality is
\begin{equation*}
  \frac{p}{1-\de_h}\left(T + Teg \right) - pg +peg+ \frac{1-p}{1-\de_h}\left(W-D\right) \geq \frac{1}{1-\de_h}\left(W-D\right)
\end{equation*}
This simplifies to
\begin{equation*}
  \frac{1}{1-\de_h}\left(T + Teg \right) - g +eg \geq \frac{1	}{1-\de_h}\left(W-D\right)
\end{equation*}
\begin{equation*}
  T + Teg - (1-\de_h)(1-e)g \geq W-D
\end{equation*}
\begin{equation*}
  T + Teg - (1-\de_h)(1-e)g \geq W-D
\end{equation*}
\begin{equation*}
  T + D - W \geq \left[Te - (1-\de_h)(1-e)\right]g 
\end{equation*}
If $Te < (1-\de_h)(1-e)$, this is a negative lower bound on the size of the concession and thus implies no additional restriction. If $Te = (1-\de_h)(1-e)$, there is again no restriction because $T+D >W$ by assumption. If $Te > (1-\de_h)(1-e)$, the mechanism is both individually rational and incentive compatible for the high type when $\frac{ T + D - W}{Te - (1-\de_h)(1-e)}>g$. \hfill $\blacksquare$\\
\\
\emph{Proof of Theorem~\ref{theorem:medwel}:}\\
An example suffices to prove each part.
	\begin{enumerate}
		\item[(a)] Examine the parameterization in Corollary~\ref{corollary:matval}. $\de_h = 0.7$ and the threshold for high types to separate through concessions when the future material value of concessions is destroyed is the same as when there is no material value, i.e. 0.5. By Corollary~\ref{corollary:matval}, the \emph{concessions separating equilibrium with inefficient gifts} is possible. Corollary~\ref{corollary:matval} also showed that the \emph{concessions separating equilibrium with efficient gifts} is not possible because the relevant patience threshold of $0.979$ is too high.

Using the mediation mechanism $M$ with $e=0.29$ and $g=12.98$, we satisfy both the low type constraint that $g\geq \frac{\left(1 - \de_l \right)\left(T +D \right)}{p - e\left(p +(1-\de_l)T +W \right)}=9.5890411$ and the high type constraint $g\leq \frac{ T + D - W}{Te - (1-\de_h)(1-e)}=12.987013$. Welfare is 5.0768, compared to 2.82 in the \emph{concessions separating equilibrium with inefficient gifts}.

		\item[(b)] Let $T = 1$, $W=1$ and $D = 1$. The assumption that $T > W - D$ is satisfied, and the threshold for high types to separate through concessions when the future material value of concessions is destroyed is $\de = \frac{1}{1+1} = .5$. Let $\de_h = \frac{15}{16}=0.9375$ and $\de_l = .5$ and $p = .5$. By Theorem~\ref{theorem:2}, separation through concessions is possible.
